---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```
# tzar

This package encapsulates R-related code for using the tzar code-running tool to manage doing lots of experiments.  tzar itself is documented at and downloadable from https://tzar-framework.atlassian.net/wiki/display/TD/Tzar+documentation.

At the moment, the code here in the tzar package is primarily the R code required for emulating running tzar so that you get the tzar setup, file naming, and parameter file building without fully running tzar.  This allows you to run your code inside of R or RStudio in a way that allows you to do debugging, which is difficult or impossible to do when tzar has control of the entire process, e.g., on a remote machine.

## Installation  
You can install tzar from github with:  
```R
# install.packages ("devtools")  
devtools::install_github ("langfob/tzar")
```

##  The basic idea behind the emulator (for experienced tzar users)

The basic idea is that running tzar with an empty model.R produces all the same setup as running with a model.R that calls your applciation code.  Consequently, you could run tzar with an empty model.R, go look for the most recent directory that tzar had created, then load the parameters.R file from that directory.  At that point, you could run your own application code with the newly loaded parameters list and it would be nearly indistinguishable from running your code under tzar but you would have control of your code for debugging, etc.  

The emulator just automatically manages the finding, loading, and running for you.  It also manages a few other details of directory naming to indicate that a job ran and/or failed under emulation.To use the emulator, you just have to add a call to one of the emulator's functions in your model.R and another one in your application code.  

## Simple example

Here is a simple example program to illustrate how to use the tzar emulator and how it works.  We will consider a trivial program to print the value of a variable x.  Not all of the example will make perfect sense until you have read the vignette, but it will give you a concrete idea of how you use tzar emulation and that should help make the details in the vignette easier to understand.

Suppose that we want to build a program that will just print the value of an input parameter x.  Call the program print_x() and suppose that it is in a file called print_x.R.  

```{r}
print_x <- function (x)
    {
    cat ("\nValue of x = '", x, "'\n\n", sep='')
    }
```

### Print a value from a parameters list.
```{r}
print_x <- function (parameters)
    {
    cat ("\nValue of x = '", parameters$x, "'\n\n", sep='')
    }

parameters <- list (x="a value from a list of parameters")
print_x (parameters)
```

### Print a value from a parameters list specified in a project.yaml file

Suppose that the project.yaml file contains:

```{yaml eval=FALSE}
project_name: simple_program
runner_class: RRunner

base_params:
    x: 100
```

#### To run under tzar emulation

We have to do 3 things:  

- First, make sure that the tzar emulator package is installed (as described above) and loaded with a call such as "library(tzar)".  
- Second, we would have to edit the **model.R** file to set the argument variables correctly.  Assuming that:  

    - the tzar jar file was in the directory "~/tzar"   
    - we want the tzar emulation scratch file in our home directory (though you can put it anywhere you want) 
        - tzar emulation requires the use of a tiny scratch file that you don't need to know anything about, other than specifying where to put it.  
        - The file is deleted at the end of the run, so its location is not terribly important.  
    - the R code is in the current working directory  
    
```{R eval=FALSE}
library (tzar)
source ("print_x.R")

main_function                 = print_x
projectPath                   =  "."
tzarEmulation_scratchFileName = "~/tzar_em_scratch.yaml"

model_with_possible_tzar_emulation (parameters,
                                    main_function,
                                    projectPath,
                                    tzarEmulation_scratchFileName
                                    )

```

- Third, we would make a **function call** something like this one at the R prompt to run the program under tzar emulation:

```{R eval=FALSE}
run_mainline_under_tzar_or_tzar_emulation (
    main_function               = print_x,    #  note, no quotes on name
    projectPath                 = ".",
    tzarJarPath                 = "~/tzar/tzar.jar", 
    emulation_scratch_file_path = "~/tzar_em_scratch.yaml",
    emulatingTzar               = TRUE
    )

```

#### To run under normal tzar without emulation

We would leave model.R as above and then just change the final argument to  run_mainline_under_tzar_or_tzar_emulation(), i.e., change emulatingTzar to FALSE.  Note that:  
- You don't have to run tzar yourself; the emulator runs it for you, and  
- This only works for local running of tzar for a single run since it's intended only to be used for development.  Once in production and spawning lots of runs, you would go back to normal command-line calls to tzar.
    - Even this could be changed though, if we were to add tzar-control arguments to the run_mainline...() call and pass them on to its call to execute tzar.

```{R eval=FALSE}
run_mainline_under_tzar_or_tzar_emulation (
    main_function               = print_x,    #  note, no quotes on name
    projectPath                 = ".",
    tzarJarPath                 = "~/tzar/tzar.jar", 
    emulation_scratch_file_path = "~/tzar_em_scratch.yaml",
    emulatingTzar               = FALSE
    )

```

##  User steps required to enable tzar emulation

For each project that uses tzar emulation, you will need to do the steps below,  once at the start of the project. There are a fair number of these steps but they are all pretty simple.  The reason for all this is that the whole process is a complete hack aimed at deceiving tzar into doing what we want.  There has been talk of adding a "dry run" option to tzar to properly do what this hack does, but so far, it's just talk.  In the meantime, this hack works.

- **Copy the model.R file** from the package into your R source code area and in it, **set the following values** of the function called **in model.R**:
    - variables to set
        - main_function
        - projectPath
        - tzarJarPath
        - emulation_scratch_file_path
    - You may be able to add other code to the model.R file, but there are no guarantees that it will still run correctly if you do.  You can examine the code in the function called there (model_with_possible_tzar_emulation()) to see whether other additions to model.R will be a problem.  Generally though, model.R only exists to call your own code and won't need anything other than what is given in the template version.
- Make sure that your **mainline application function has just one argument, a _parameters_ argument**, regardless of whether it's used in your code or not.
- Make sure that the **tzar package is loaded** for your code, e.g., "library(tzar)".
- Make sure that you have a **project.yaml file in the projectPath directory**.   
    - This is a tzar requirement and not specific to tzar emulation.
- **When running emulation** rather than normal tzar, make sure that the **project.yaml file is only doing a single run** rather than using tzar's ability to generate lots of runs (e.g., with a repetitions section).  
    - If you were to generate multiple runs, there would be ambiguity because there would be more than one place to look for the parameters.R file that tzar generates.
- **Call the following function** at the R prompt or on RStudio's command line to invoke the whole process, i.e., to run your mainline application code under tzar or tzar emulation:

```{r eval=FALSE}
run_mainline_under_tzar_or_tzar_emulation (main_function,
                                           projectPath,
                                           tzarJarPath, 
                                           emulation_scratch_file_path,
                                           emulatingTzar
                                           )
```

