---
title: "How to use the tzar package"
author: "Bill Langford"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

##  Overview

The java program tzar helps with managing many runs of sets of experiments and making them easier to reproduce.  However, one problem with tzar is that it runs your R code in a separate process that makes it difficult, if not impossible, to easily debug your R code.  

The tzar package is intended to make it easy to emulate running tzar while keeping the running code under your control (e.g., inside RStudio).  This allows more interaction with your code, e.g., allowing the use of the browser() function to pause and examine variables.  Similarly, when your program crashes in emulation mode you can get at lots of information that is lost or truncated when the program is run in a separate tzar process.  

## Simple example _[move to README?]_

Here, we will give a simple example program to illustrate how to use the tzar emulator and how it works.  We will consider a trivial program to print the value of a variable x.  Not all of the example will make perfect sense until you have read the rest of this vignette, but it will give you a concrete idea of how you use tzar emulation and that should help make the later details easier to understand.

Suppose that we want to build a program that will just print the value of an input parameter x.  Call the program print_x() and suppose that it is in a file called print_x.R.  

```{r}
print_x <- function (x)
    {
    cat ("\nValue of x = '", x, "'\n\n", sep='')
    }
```

### Print a value from a parameters list.
```{r}
print_x <- function (parameters)
    {
    cat ("\nValue of x = '", parameters$x, "'\n\n", sep='')
    }

parameters <- list (x="a value from a list of parameters")
print_x (parameters)
```

### Print a value from a parameters list specified in a project.yaml file

Suppose that the project.yaml file contains:

```{yaml eval=FALSE}
project_name: simple_program
runner_class: RRunner

base_params:
    x: 100
```

#### To run under tzar emulation

We have to do 2 things:  

- First, we would have to edit the **model.R** file to set the argument variables correctly.  Assuming that:  

    - the tzar jar file was in the directory "~/tzar"   
    - we want the tzar emulation scratch file in our home directory  
        - tzar emulation requires the use of a tiny scratch file that you don't need to know anything about, other than specifying where to put it.  
        - The file is deleted at the end of the run, so its location is not terribly important.  
    - the R code is in the current working directory  
    
```{R eval=FALSE}
library (tzar)
source ("print_x.R")

main_function                 = print_x
projectPath                   =  "."
tzarEmulation_scratchFileName = "~/tzar_em_scratch.yaml"

model_with_possible_tzar_emulation (parameters,
                                    main_function,
                                    projectPath,
                                    tzarEmulation_scratchFileName
                                    )

```

- Second, we would make a **function call** something like this one at the R prompt to run the program under tzar emulation:

```{R eval=FALSE}
run_mainline_under_tzar_or_tzar_emulation (
    main_function               = print_x,    #  note, no quotes on name
    projectPath                 = ".",
    tzarJarPath                 = "~/tzar/tzar.jar", 
    emulation_scratch_file_path = "~/tzar_em_scratch.yaml",
    emulatingTzar               = TRUE
    )

```

#### To run under normal tzar without emulation

We would leave model.R as above and then just change the final argument to  run_mainline_under_tzar_or_tzar_emulation(), i.e., change emulatingTzar to FALSE:

```{R eval=FALSE}
run_mainline_under_tzar_or_tzar_emulation (
    main_function               = print_x,    #  note, no quotes on name
    projectPath                 = ".",
    tzarJarPath                 = "~/tzar/tzar.jar", 
    emulation_scratch_file_path = "~/tzar_em_scratch.yaml",
    emulatingTzar               = FALSE
    )

```

##  How tzar emulation works

Basically, tzar emulation is just a way to trick tzar into giving you half of its benefits.  The various things that you have to do as a user to use the package are largely a consequence of how what is required to trick tzar.  This section gives some background that helps understand what's necessary to trick tzar.

#### Conventions used by tzar that matter in emulation

- **project.yaml**:  tzar requires a project.yaml file to exist in the directory where the source code is stored.
    - tzar builds a *parameters* list by decoding the yaml expressions in the project.yaml file.  
    - This *parameters* list is then written out to a file "parameters.R" which can be read back in as a list object by any R program.
    - One useful thing about tzar is that it builds various directory locations and variables for you by replacing wildcard values, but the names that are built aren't known to you until after tzar has started up.  
        - However, the values of these constructed names and variables are added to the *parameters* list before it is written out to "parameters.R".
        - After tzar builds the directory path to its output, it saves it in the *parameters* list as parameters$fullOutputDirWithSlash.  
            - This is the location where the *parameters* list itself will be written as a parameters.R file.
- **model.R**:  tzar requires you to have a model.R file (in the same directory) that invokes the code you want to run.
    - The code in **model.R** is run *after* tzar has loaded the project.yaml data into the *parameters* list.
    - This means that model.R can treat the variable *parameters* as a pre-existing variable inside the same scope as the code in model.R.
    - The code in model.R has no restrictions on what it can or has to do.  It could be completely empty for all that tzar cares.  
        - Consequently, we can run a model.R that does almost nothing at all but it will still have read and written the *parameters* list.
    
#### How the conventions suggest a method for tzar emulation

Tzar creates the parameters list and saves it to disk *regardless of what is done in model.R*.  So, in emulation mode we can take advantage of this by having model.R do nothing at all, except save the path to the directory where the parameters list was written.  

We can easily do this by just having model.R write that path into a scratch file whenever it's doing emulation. The emulation code can then read the location from the scratch file and find the "parameters.R" file to load the parameters list and pass the reconstituted  *parameters* to our mainline application code.  This parameters list will contain all of the paths and variable names that tzar has built based on the project.yaml file so that we don't have to build them ourselves.
    
####  Allowing both normal tzar and tzar emulation

The one other thing that we would like is for our code to still work with as little change as possible when running normally under tzar rather than emulating tzar.  This is easy to accomplish by just adding a boolean flag indicating whether we are running tzar normally or just emulating it.  Each time you want to switch from emulation to normal tzar and vice versa, you just have to change the value of this flag.  

The only hitch is that the flag has to be known to both model.R and the emulation code.  At the moment, the simplest way to make sure that the value is the same in both bits of code is to read it from a file that both bits of code know about. This is a little bit clumsy, but it works.

##  Implementation summary

Implementation of tzar emulation requires two pieces of code:  

* One piece that runs inside model.R whenever tzar is invoked, and 
* Another piece that wraps around our mainline application code to: 
    + run tzar, and 
    + get the *parameters* list back from it if running in emulation mode, and
    + run the mainline application.  
    
In pseudocode, this looks like the following:

### In project.yaml:
```
Set emulatingTzar flag to TRUE if emulating, FALSE if running normal tzar.
```

### In wrapper R code around main application:
```
run tzar

if (parameters$emulating tzar)  
    read tzar output directory path from scratch file
    load parameters list from tzar output directory
    run mainline application
```

###  In model.R:

```
Assume called by tzar and tzar has already loaded parameters list from project.yaml

if (parameters$emulatingTzar)  
    do nothing except write tzar output directory path to scratch file
else  
    run mainline application
```

##  User steps required to enable tzar emulation _[move to README?]_

For each project that uses tzar emulation, you will need to do the steps below,  once at the start of the project. There are a fair number of these steps but they are all pretty simple.  The reason for all this is that the whole process is a complete hack aimed at deceiving tzar into doing what we want.  There has been talk of adding a "dry run" option to tzar to properly do what this hack does, but so far, it's just talk.  In the meantime, this hack works.

- **Copy the model.R file** from the package into your R source code area and in it, **set the following values** of the function called **in model.R**:
    - variables to set
        - main_function
        - projectPath
        - tzarJarPath
        - emulation_scratch_file_path
    - You may be able to add other code to the model.R file, but there are no guarantees that it will still run correctly if you do.  You can examine the code in the function called there (model_with_possible_tzar_emulation()) to see whether other additions to model.R will be a problem.  Generally though, model.R only exists to call your own code and won't need anything other than what is given in the template version.
- Make sure that your **mainline application function has just one argument, a _parameters_ argument**, regardless of whether it's used or not in your code or not.
- Make sure that the **tzar package is loaded** for your code, e.g., "library(tzar)".
- Make sure that you have a **project.yaml file in the projectPath directory**.   
    - This is a tzar requirement and not specific to tzar emulation.
- **When running emulation** rather than normal tzar, make sure that the **project.yaml file is only doing a single run** rather than using tzar's ability to generate lots of runs (e.g., with a repetitions section).  
    - If you were to generate multiple runs, there would be ambiguity because there would be more than one place to look for the parameters.R file that tzar generates.
- **Call the following function** at the R prompt or on RStudio's command line to invoke the whole process, i.e., to run your mainline application code under tzar or tzar emulation:

```{r eval=FALSE}
run_mainline_under_tzar_or_tzar_emulation (main_function,
                                           projectPath,
                                           tzarJarPath, 
                                           emulation_scratch_file_path,
                                           emulatingTzar
                                           )
```


---
---


# Extraneous material that will probably end up removed from this vignette

##  NOTES

- Trying to run the following call, I have various problems:

```{r eval=FALSE}
run_mainline_under_tzar_or_tzar_emulation (
         main_function=trial_main_function,
         projectPath=".",
         tzarJarPath = "~/D/rdv-framework-latest-work/tzar.jar",
         emulation_scratch_file_path="~/tzar_emulation_scratch.yaml",
         emulatingTzar=TRUE
         )
```

- Can't get the code to work inside the testthat code when running CHECK.  
    - I have no idea what's wrong there.
- IMPORTANT:  Running the code *outside* of CHECK only works if you first:
    - cd into the tzar/tests/testthat directory and 
    - source ("model_support_code.R")



##  Some useful tzar-related commands for reference

```
java   -jar   tzar.jar   scheduleruns   http://rdv-framework.googlecode.com/svn/trunk/projects/rdvPackages/biodivprobgen/R    --clustername   biodivprobgen   --revision   head   --runset   biodivprobgen_1kruns_1sce_8001ranSeed_10to60ranPUs
```

```
java -jar tzar.jar execlocalruns /Users/bill/D/rdv-framework/projects/rdvPackages/biodivprobgen/R/ --runset biodivprobgen_1kruns_1sce_8001ranSeed_10to60ranPUs
```

```
java -jar tzar.jar execlocalruns https://tzar-framework.googlecode.com/svn/tags/latest/example-projects/example-R
```

```
https://tzar-framework.googlecode.com/svn/trunk/
```
```
sudo  scp  -i  /Users/bill/.ssh/rdv.pem  /Users/bill/D/rdv-framework/tzar-0.5.5-runset-wildcard.mar.07.2015.jar  ubuntu@146.118.96.116:/home/ubuntu
```
```
 [ubuntu@unixnpeap06 tzar_output]$ grep -i bdpg ?????_scen_rand_10_to_60_PUs_0_5_optFrac_random_r_and_p/prob_diff_results.csv > ../aggregated1kResults.csv
-bash: ../aggregated1kResults.csv: Read-only file system
```

===

Vignettes are long form documentation commonly included in packages. Because they are part of the distribution of the package, they need to be as compact as possible. The `html_vignette` output type provides a custom style sheet (and tweaks some options) to ensure that the resulting html is as small as possible. The `html_vignette` format:

- Never uses retina figures
- Has a smaller default figure size
- Uses a custom CSS stylesheet instead of the default Twitter Bootstrap style

## Vignette Info

Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
