---
title: "How to use the tzar package"
author: "Bill Langford"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The java program tzar helps with managing many runs of sets of experiments and making them easier to reproduce.  One problem with tzar is that it runs your R code in a separate process that makes it difficult, if not impossible, to easily debug your R code.  

The tzar package is intended to make it easy to emulate running tzar while keeping the running code under your control to allow more interaction with your code, e.g., allowing the use of the browser() function to pause and examine variables.  Similarly, when your program crashes in emulation mode you can get at lots of information that is lost when the program is run in a separate tzar process.  

Here, we will give a detailed, incremental buildup of a small example program to illustrate how to use the tzar emulator and how it works.  We will consider a trivial program to print the value of a variable x.  We will walk through a common sequence of steps in the evolution of a program and show how to this might happen in the context of tzar and the tzar emulator.

## Simple program

Suppose that we want to build a program that will just print the value of an input parameter x.  Call the program print_x().  

```{r}
print_x <- function (x)
    {
    cat ("\nValue of x = '", x, "'\n\n", sep='')
    }
```

#### Print a value from a parameters list.
```{r}
parameters <- list (x="a value from a list of parameters")
print_x (parameters$x)
```

##  Program using tzar and tzar emulation

#### Conventions used by tzar that matter in emulation

- tzar builds a *parameters* list by reading a project.yaml file from the directory where the program source code is stored.  
    - This *parameters* list is then written out to a file parameters.R which can be read back in as a list object by any R program.
- tzar requires you to have a model.R file (in the same directory) that invokes the code you want to run.
    - The code in model.R is run *after* tzar has loaded the project.yaml data into the *parameters* list.
    - This means that model.R can treat the variable *parameters* as a pre-existing variable inside the same scope as the code in model.R.
    - The code in model.R has no restrictions on what it can or has to do.  It could be completely empty for all that tzar cares.  
        - Consequently, we can run a model.R that does almost nothing at all but it will still have read and written the *parameters* list.
- One useful thing about tzar is that it builds various directory locations and variables for you, but the names that are built aren't known to you until after tzar has started up.  
    - However, the values of these constructed names and variables are added to the *parameters* list.
    - After tzar builds the directory path to its output, it saves it in the *parameters* list as parameters$fullOutputDirWithSlash.  
        - This is the location where the *parameters* list itself will be written as a parameters.R file.
    
#### How the conventions suggest a method for tzar emulation

Tzar creates the parameters list and saves it to disk *regardless of what is done in model.R*.  So, in emulation mode we can take advantage of this by having model.R do nothing at all, except save the path to the directory where the parameters list was written.  We can easily do this by just having model.R write that path into a scratch file whenever it's doing emulation. The emulation code can then read the location from the scratch file and find the parameters.R file to load the parameters list and pass the reconstituted  *parameters* to our mainline application code.  This parameters list will contain all of the paths and variable names that tzar has built based on the project.yaml file so that we don't have to build them ourselves.
    
####  Allowing both normal tzar and tzar emulation

The one other thing that we would like is for our code to still work with as little change as possible when running under tzar rather than emulating tzar.  This is easy to accomplish by just adding a boolean flag indicating whether we are running tzar normally or just emulating it.  Each time you want to switch from emulation to normal tzar and vice versa, you just have to change the value of this flag.  

The only hitch is that the flag has to be known to both model.R and the emulation code.  At the moment, the simplest way to make sure that the value is the same in both bits of code is to read it from a file that both bits of code know about. This is a little bit clumsy, but the flag's value will probably only be changed very rarely, i.e., when you first start the project, emulation will be on and then when things are working well enough to do final experiments, you switch emulation off by editing this file.

##  Summary

So, to summarize, our implementation of tzar emulation requires two pieces of code:  

* one piece that runs inside model.R whenever tzar is invoked and 
* another piece that wraps around our mainline application code to: 
    + run tzar and 
    + get the *parameters* list back from it if running in emulation mode
    + run the mainline application.  
    
In pseudocode, this looks like the following:

### In project.yaml:
```
Set emulatingTzar flag to TRUE if emulating, FALSE if running normal tzar.
```

### In wrapper R code around main application:
```
run tzar

if (parameters$emulating tzar)  
    read tzar output directory path from scratch file
    load parameters list from tzar output directory
    run mainline application
```

###  In model.R:

```
Assume called by tzar and tzar has already loaded parameters list from project.yaml

if (parameters$emulatingTzar)  
    do nothing except write tzar output directory path to scratch file
else  
    run mainline application
```

##  User steps required to enable tzar emulation

For each project that uses tzar emulation, you will need to do the following once  at the start of the project.  You may also need to change the *emulatingTzar* flag at some time during the project when you switch from emulation to running tzar normally.

1. Copy the model.R file given here into your R source code area and in it, set the following values of the function called in model.R:

    - main_function
    - projectPath
    - tzarJarPath

2. Make sure that your mainline application function has the following 3 arguments, regardless of whether they're used or not:

    - parameters
    - running_tzar_or_tzar_emulator
    - emulatingTzar
3. Make sure that build_scratch_file_name() is writing the emulation scratch file where you want it.
4. Make sure that the tzarEmulation flag file exists in the appropriate directory and contains either TRUE or FALSE and nothing else.
5. Make sure that the tzar package is loaded for your code.
6. When running emulation rather than normal tzar, make sure that the project.yaml file is only doing a single run rather than using tzar's ability to generate lots of runs.
7. Call the following function to run your mainline application code under tzar or tzar emulation:

```{r eval=FALSE}
run_mainline_under_tzar_or_tzar_emulation (parameters,
                                               main_function,
                                                projectPath,
                                                tzarJarPath,                                                running_tzar_or_tzar_emulator
                                               )
```

##  NOTES

- Trying to run the following call, I have various problems:

```{r eval=FALSE}
run_mainline_under_tzar_or_tzar_emulation (main_function=trial_main_function,
         projectPath=".",
         tzarJarPath = "~/D/rdv-framework-latest-work/tzar.jar",
         emulation_scratch_file_path="~/tzar_emulation_scratch.yaml",
         emulatingTzar=TRUE
         )
```

- Can't get the code to work inside the testthat code when running CHECK.  
    - I have no idea what's wrong there.
- IMPORTANT:  Running the code *outside* of CHECK also fails if you don't first:
    - cd into the tzar/tests/testthat directory and 
    - source ("model_support_code.R")



##  Some useful tzar-related commands for reference

```
java   -jar   tzar.jar   scheduleruns   http://rdv-framework.googlecode.com/svn/trunk/projects/rdvPackages/biodivprobgen/R    --clustername   biodivprobgen   --revision   head   --runset   biodivprobgen_1kruns_1sce_8001ranSeed_10to60ranPUs
```

```
java -jar tzar.jar execlocalruns /Users/bill/D/rdv-framework/projects/rdvPackages/biodivprobgen/R/ --runset biodivprobgen_1kruns_1sce_8001ranSeed_10to60ranPUs
```

```
java -jar tzar.jar execlocalruns https://tzar-framework.googlecode.com/svn/tags/latest/example-projects/example-R
```

```
https://tzar-framework.googlecode.com/svn/trunk/
```
```
sudo  scp  -i  /Users/bill/.ssh/rdv.pem  /Users/bill/D/rdv-framework/tzar-0.5.5-runset-wildcard.mar.07.2015.jar  ubuntu@146.118.96.116:/home/ubuntu
```
```
 [ubuntu@unixnpeap06 tzar_output]$ grep -i bdpg ?????_scen_rand_10_to_60_PUs_0_5_optFrac_random_r_and_p/prob_diff_results.csv > ../aggregated1kResults.csv
-bash: ../aggregated1kResults.csv: Read-only file system
```

===

Vignettes are long form documentation commonly included in packages. Because they are part of the distribution of the package, they need to be as compact as possible. The `html_vignette` output type provides a custom style sheet (and tweaks some options) to ensure that the resulting html is as small as possible. The `html_vignette` format:

- Never uses retina figures
- Has a smaller default figure size
- Uses a custom CSS stylesheet instead of the default Twitter Bootstrap style

## Vignette Info

Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
